---
output:
  word_document: default
  html_document: default
---
Single Continuos Variable: (dataset Chickwts)

* 1. Histogram - hist()

* 2. Density plot - plot()

* 3. Box-Whisker Plot - boxplot()


Single Discrete Variable: (dataset chickwts)

* 4. Bar Chart - barplot()


Two Continuous Variable : (dataset cars)

* 5. Scatter Plot - plot()


Two Variable: One Continuous, One Discrete (Dataset chickwts)

* 6. Box-Whisker Plot - boxplot() - see 3 above as well

* 7. Pie Chart - pie()

* 8. Dot Chart - dotchart()

* 9. Strip Chart - stripchart()


Two Variables: Both Discrete (Dataset ggplot2::mpg)

* 10. Mosaic Plot - mosaicplot()

* 11 Stacked Bar Plot - barplot()

Time series: (Dataset LakeHuron)

* 12. Line Charts - plot()


Understanding a Single Continuous Numeric Variable
  
###1. Histogram

```{r}
data("chickwts")
str(chickwts)
```
```{r}
summary(chickwts)
```

```{r}
head(chickwts)
```

```{r}
hist(chickwts$weight)
```

```{r}
hist(chickwts$weight,xlab = "Weight",ylab = "Frequency",
     main = "Chicken Weight",breaks = 20)
```

```{r}
hist(chickwts$weight, breaks = "FD")
```

```{r}
bin_width <- 5
breaks <- seq(min(chickwts$weight), max(chickwts$weight), by = bin_width)
hist(chickwts$weight, breaks = breaks)
```

```{r}
hist(chickwts$weight,
     xlab = "Weight",
     ylab = "Frequency",
     main = "Chicken Weight",
     breaks = c(100,200,300,400,500))
```
```{r}
hist(chickwts$weight,
     xlab = "Weight",
     ylab = "Frequency",
     main = "Chicken Weight",
     breaks = c(100,200,300,400,500),
     labels = T,
     col = c("red","green","blue","yellow"))
```
### 2. Density Plot for single continuous variable

```{r}
data = chickwts$weight
den = density(data)
plot(den)
```

```{r}
plot(den, 
     main = "Density Plot with Filled Area", 
     xlab = "Chicken Weight", 
     ylab = "Density", 
     col = "blue", 
     lwd = 2)
```
Box plot for single continuous variable

```{r}
boxplot(chickwts$weight)
```

```{r}
boxplot(chickwts$weight,horizontal = T)
```

```{r}
boxplot(chickwts$weight,
        notch = T,
        main = "Box PLot for \n Weight of Chicken",
        xlab = "Weight of Chicken",
        ylab = "Weights",
        col = "lightblue",
        border = "black",
        outpch = 16,
        outcol = "red")
```
##plotting single discrete varible

#Bar plot for single discrete varible

```{r}
bp = table(chickwts$feed)
barplot(bp)
```
```{r}
bp = barplot(bp,
        col = rainbow(nlevels(chickwts$feed)),
        xlab = "Feed of Category",
        ylab = "Count",
        main = "Bar PLot for Feed of Chickweights")
```
##Plotting two continuous variable

#Scatter Plot

```{r}
data("cars")

plot(cars$speed,cars$dist)
```
```{r}
plot(jitter(cars$speed),jitter(cars$dist))
```
```{r}
plot(jitter(cars$speed),jitter(cars$dist),xlab = "Speed of Cars",
     ylab = "Stopping Distance according to speed",
     main = "Distance to Stop",
     pch = 4)
```

```{r}
plot(jitter(cars$speed),jitter(cars$dist),xlab = "Speed of Cars",
     ylab = "Stopping Distance according to speed",
     main = "Distance to Stop",
     pch = 4)
abline(a = 60,b = 0,v = 15,h = 20)
abline(lm(cars$dist ~ cars$speed))
text(x=10,y=100,label = "Tested on the 1st of January")
```

##Two variable one continuous and one variable

#Boxplot

```{r}
boxplot(chickwts$weight ~ chickwts$feed)
```

#Pie Chart

```{r}
defects = c(9,15,42,29)
names(defects) = c("A","B","C","D")

pie(defects)
```
```{r}
lab = c("A","B","C","D")
pie(defects,labels = lab,col = c("red","yellow","green","blue"))
```
#Dot Chart

```{r}
dotchart(chickwts$weight,labels = chickwts$feed)
```

```{r}
dotchart(chickwts$weight,group = chickwts$feed)
```
#Strip Chart

```{r}
stripchart(chickwts$weight ~ chickwts$feed)
```
#Stacked bar chart

```{r}
cy = table(mtcars$vs,mtcars$cyl)
barplot(cy,legend =rownames(cy))
```

##Latice packages

##Histogram


```{r}
library(lattice)
```

```{r}
histogram(~weight, data = chickwts)
```

```{r}
histogram(~weight | factor(feed),data = chickwts)
```

```{r}
histogram(~mpg, data = mtcars)
```

```{r}
histogram(~mpg | factor(cyl),data = mtcars)
```
```{r}
histogram(~mpg | factor(am),data = mtcars)
```
#If the factor is continuous

```{r}
histogram(~mpg | wt, data = mtcars)
```
```{r}
spli = equal.count(mtcars$wt,3)
spli
```

```{r}
histogram(~mpg | spli, data = mtcars)
```

```{r}
histogram(~mpg | spli, data = mtcars, layout = c(1,3))
```
#Density plot

```{r}
chden = density(chickwts$weight)
chden
```

```{r}
plot(chden)
```
```{r}
densityplot(~weight, data = chickwts)
```
```{r}
densityplot(~weight, data = chickwts,plot.points = F)
```
```{r}
densityplot(~weight | feed , data = chickwts)
```
```{r}
data("mtcars")
```

```{r}
densityplot(~mpg | factor(cyl), data = mtcars)
```
```{r}
densityplot(~mpg | factor(cyl),data = mtcars, layout = c(1,3))
```

```{r}
densityplot(~mpg | factor(cyl), data = mtcars, layout = c(1,3),
            groups = factor(am))
```
```{r}
densityplot(~mpg | factor(cyl),data = mtcars, layout = c(1,3),
            groups = factor(am),
            auto.key = T)
```
#Boxplot

```{r}
boxplot(chickwts$weight)
```
```{r}
boxplot(chickwts$weight ~ chickwts$feed)
```
```{r}
bwplot(weight ~ feed, data = chickwts)
```

```{r}
bwplot(~weight | factor(feed), data = chickwts)
```
```{r}
bwplot(~weight | factor(feed), data = chickwts, notch = T)
```
```{r}
bwplot(~weight | factor(feed), data = chickwts, notch = T,
       layout = c(1,6))
```
```{r}
bwplot(weight ~ feed , data = chickwts,panel = panel.violin)
```
#Barchart


```{r}
tb = table(chickwts$feed)
tb
```
```{r}
barplot(tb)
```
```{r}
barchart(~tb,data = chickwts)
```
```{r}
barchart(rownames(mtcars) ~ mpg, data = mtcars)
```

```{r}
barchart(rownames(mtcars) ~ mpg | factor(gear) + factor(am)
           ,data = mtcars)
```
#Scatter Plot

```{r}
xyplot(mpg ~ wt, data = mtcars)
```
```{r}
xyplot(mpg ~ wt, data = mtcars, grid = T)
```

```{r}
xyplot(mpg ~ wt, data = mtcars, type = c("p","r"))
```
```{r}
xyplot(mpg ~ wt, data = mtcars, type = c("p","smooth"))
```

```{r}
xyplot(mpg ~ wt | factor(gear), data = mtcars, type = c("p","smooth"),
       grid = T)
```

```{r}
xyplot(mpg ~ wt | factor(gear), data = mtcars,
       type = c("p","smooth"),
       grid = T,
       col.line = "red",lwd = 3)
```

```{r}
xyplot(mpg ~ wt | factor(gear), data = mtcars,
       type = c("p","smooth"),
       grid = T,
       groups = factor(am),auto.key = T)
```
#Scater Plot Matrix

```{r}
splom(chickwts)
```
```{r}
splom(mtcars[c(1,2,3,4,6)], type = c("p","smooth"))
```

```{r}
cloud(mpg ~ disp*wt, data = mtcars)
```
```{r}
cloud(mpg ~ disp*wt, data = mtcars, groups = factor(cyl),
      auto.key = T)
```
#Dot and strip plot

```{r}
dotchart(chickwts$weight, groups = chickwts$feed)
```

```{r}
dotplot(~weight | feed, data = chickwts)
```

##Visualization using ggplot2 package

```{r}
library(ggplot2)
```

```{r}
ggplot(data = mtcars, aes(x=mpg))+geom_histogram()
```
```{r}
ggplot(data = mtcars,aes(x=mpg))+geom_histogram(binwidth = 6)
```

```{r}
ggplot(data = mtcars,aes(x=mpg))+geom_density()
```
```{r}
ggplot(data = mtcars,aes(x=mpg))+geom_density()
```
```{r}
ggplot(data = mtcars,aes(x=mpg))+geom_dotplot()
```

```{r}
ggplot(data = mtcars,aes(mpg))+geom_area(stat = "bin")
```
```{r}
ggplot(data = mtcars,aes(mpg))+geom_area(stat = "bin",bins = 6)
```
```{r}
ggplot(data = mtcars,aes(mpg))+geom_freqpoly()
```
```{r}
ggplot(data = mtcars,aes(mpg))+geom_density(color = "blue")
```
```{r}
ggplot(data = mtcars,aes(mpg))+geom_density(color = "blue",
                                             fill = "red",
                                             alpha = 0.5)
```

```{r}
ggplot(data = mtcars,aes(mpg))+geom_density(color = "blue",
                                            fill = "red",
                                            alpha = 0.5,
                                            size = 3,
                                            linetype = "dotted")
```
```{r}
data("diamonds")
```

```{r}
ggplot(data = diamonds,aes(x=carat,y=price,
                           shape = cut,colour = clarity))+
  geom_point()
```
```{r}
ggplot(data = diamonds,aes(x=carat,y=price,
                           shape = cut,colour = clarity))+
  geom_point()+geom_smooth()
```

```{r}
d2 = diamonds[diamonds$carat <= 0.31 & diamonds$carat >= 0.29,]
d2
```

```{r}
ggplot(data = d2,aes(x=cut,y=price,colour = clarity))+
  geom_boxplot()
```
```{r}
ggplot(data = d2,aes(x=cut,y=price,fill = clarity))+
  geom_boxplot()
```

```{r}
ggplot(data = d2,aes(x=clarity))+geom_bar()
```

```{r}
ggplot(data = d2,aes(x=clarity,fill = cut))+geom_bar()
```
```{r}
ggplot(data = d2,aes(x=clarity,fill = cut))+
  geom_bar(position = "fill")
```

```{r}
ggplot(data = d2,aes(x=clarity,fill = cut))+
  geom_bar(position = "dodge")
```

```{r}
ggplot(data = d2,aes(x=clarity,fill = cut))+
  geom_bar(position = "stack")
```
#Line plot

```{r}
data("LakeHuron")
head(LakeHuron)
```
```{r}
lh = data.frame(year = 1875:1972,level = as.vector(LakeHuron))
head(lh)
```

```{r}
ggplot(data = lh,aes(x=year,y=level))+geom_line()
```
##Scatter plot

```{r}
ggplot(data = diamonds,aes(x=carat,y=price))+geom_point()
```
```{r}
ggplot(data = diamonds,aes(x=carat,y=price))+geom_point()+
  facet_grid(.~cut)
```
```{r}
ggplot(data = diamonds,aes(x=carat,y=price))+geom_point()+
  facet_grid(cut~.)
```
```{r}
ggplot(data = diamonds,aes(x=carat,y=price))+geom_point()+
  facet_grid(cut~clarity)
```

```{r}
ggplot(data = diamonds,aes(x=carat,y=price))+geom_point()+
  facet_wrap(~cut)
```

```{r}
ggplot(data = diamonds,aes(x=carat,y=price))+geom_point()+
  facet_wrap(~clarity)
```
```{r}
ggplot(data = diamonds,aes(x=carat,y=price))+geom_point()+
  geom_smooth(se = F)
```

```{r}
ggplot(data = diamonds,aes(x=carat,y=price))+geom_point()+
  geom_smooth(se = F,method = "lm")
```

```{r}
ggplot(data = diamonds,aes(x=carat,y=price))+geom_point()+
  stat_smooth(method = "lm",se = F)
```
#zoom

```{r}
ggplot(data = diamonds,aes(x=carat,y=price))+geom_point()+
  facet_grid(cut~color)+geom_smooth()
```
```{r}
ggplot(data = diamonds,aes(x=carat,y=price))+geom_point()+
  geom_smooth() + xlim(c(3,5))
```

```{r}
ggplot(data = diamonds,aes(x=carat,y=price))+geom_point()+
  geom_smooth() + xlim(c(3,5))
```

```{r}
ggplot(data = diamonds,aes(x=carat,y=price))+geom_point()+
  geom_smooth() + coord_cartesian(xlim = c(3,5))
```
```{r}
ggplot(data = diamonds,aes(x=carat,y=price))+geom_point()+
  geom_smooth() + coord_cartesian(xlim = c(3,5),ylim = c(5000,20000))

```

```{r}
ggplot(data = mtcars,aes(x=mpg,y=wt))+geom_point()
```

```{r}
ggplot(data = mtcars,aes(x=mpg,y=wt))+geom_point()+
  coord_fixed(ratio = 1)
```

```{r}
ggplot(data = mtcars,aes(x=mpg,y=wt))+geom_point()+
  coord_fixed(ratio = 4)
```

```{r}
ggplot(data = mtcars,aes(x=mpg,y=wt))+geom_point()+
  coord_flip()
```
#Pie chart

```{r}
cyl_summary <- as.data.frame(table(mtcars$cyl))
colnames(cyl_summary) <- c("Cylinders", "Count")
```

```{r}
ggplot(cyl_summary, aes(x = "", y = Count, fill = factor(Cylinders))) +
  geom_bar(stat = "identity", width = 1) + 
  coord_polar("y", start = 0) + 
  labs(title = "Distribution of Cylinders in mtcars",
       fill = "Cylinders") +
  theme_void()
```
#Transform Scale

```{r}
ggplot(diamonds,aes(carat,price))+geom_point()+
  scale_x_log10()+
  scale_y_log10()
```

```{r}
ggplot(diamonds,aes(carat,price))+geom_point()+
  coord_trans(x = "log10",y="log10")
```
```{r}
ggplot(diamonds,aes(carat,price))+geom_point()+
  coord_trans(x="sqrt",y="sqrt")
```
#Theme

```{r}
ggplot(diamonds,aes(carat,price))+geom_point()
```
```{r}
ggplot(diamonds,aes(carat,price))+geom_point()+theme_light()
```
```{r}
ggplot(diamonds,aes(carat,price))+geom_point()+theme_dark()
```

```{r}
ggplot(diamonds,aes(carat,price))+geom_point()+theme_gray()

```
```{r}
library(ggthemes)
```

```{r}
ggplot(diamonds,aes(carat,price))+geom_point()+theme_calc()
```

```{r}
ggplot(diamonds,aes(carat,price))+geom_point()+theme_economist()
```

```{r}
ggplot(diamonds,aes(carat,price))+geom_point()+
  theme_economist()+labs(title = "Diamonds ~ Price by carat",
                         y = "Price $",x="Carat")
```

#qplot automatically picks the right geom

```{r}
qplot(x=mpg,data = mtcars)
```

```{r}
qplot(mpg,disp,data = mtcars)
```

```{r}
qplot(x=mpg,data = mtcars,geom = "density")
```

```{r}
qplot(x=mpg,data = mtcars,geom = "histogram")
```

```{r}
qplot(x=factor(gear),y=mpg,data = mtcars,geom = "boxplot")
```
```{r}
qplot(x=factor(gear),y=mpg,data = mtcars,geom = c("point","boxplot"))
```

```{r}
qplot(x=factor(gear),y=mpg,data = mtcars,geom = c("boxplot","point"))
```

```{r}
qplot(x=factor(gear),y=mpg,data = mtcars,geom = c("boxplot","jitter"))
```

```{r}
qplot(cyl,mpg,data = mtcars,color = factor(am))
```
```{r}
qplot(cyl,mpg,data = mtcars,color = factor(am),size = wt)
```
```{r}
qplot(cyl,mpg,data = mtcars,color = factor(am),facets = gear ~ vs)

```
```{r}
qplot(cyl,mpg,data = mtcars,color = factor(am),
      main = "MPG by Number of Cylinders",
      xlab = "Number of Cylinders",
      ylab = "Miles per Gallon",
      ylim = c(0,40))
```
```{r}
qplot(cyl,mpg,data = mtcars,color = factor(am),
      main = "MPG by Number of Cylinders",
      xlab = "Number of Cylinders",
      ylab = "Miles per Gallon",
      ylim = c(0,40))+theme_bw()
```

```{r}
qplot(cyl,mpg,data = mtcars,color = factor(am),
      main = "MPG by Number of Cylinders",
      xlab = "Number of Cylinders",
      ylab = "Miles per Gallon",
      ylim = c(0,40))+theme_classic()

```
```{r}
qplot(cyl,mpg,data = mtcars,color = factor(am),
      main = "MPG by Number of Cylinders",
      xlab = "Number of Cylinders",
      ylab = "Miles per Gallon",
      ylim = c(0,40))+theme_gray()
```

